---
title: "Modelling from R for Datascience"
output: html_document
---

```{r}
library(modelr)
library(tidyverse)

```

Overall point, modeling is trying to select the best paraments to model
or "predict" one variable based on the value(s) of other(s)

What is the relationship between x and y:
```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point()
```

a very simple mode is in the form:

`y = a_0 + a_1 * x`  
different models from that family look like:
```{r}
models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)

ggplot(sim1, aes(x, y)) + 
  geom_abline(aes(intercept = a1, slope = a2), data = models, alpha = 1/4) +
  geom_point() 
```

But what is the "Best" model- the best model is one that minizes the distance between the predicted value and the actual values
```{r}
dist1 <- sim1 %>% 
  mutate(
    dodge = rep(c(-1, 0, 1) / 20, 10),
    x1 = x + dodge,
    pred = 7 + x1 * 1.5
  )
ggplot(dist1, aes(x1, y)) + 
  geom_abline(intercept = 7, slope = 1.5, colour = "grey40") +
  geom_point(colour = "grey40") +
  geom_linerange(aes(ymin = y, ymax = pred), colour = "#3366FF") 
```
We can measure the distance for random models
```{r}
model1 <- function(a, data) {
  a[1] + data$x * a[2]
}
model1(c(7, 1.5), sim1)
```
```{r}
measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff ^ 2))
}
measure_distance(c(7, 1.5), sim1)
#> [1] 2.665212
```
```{r}
sim1_dist <- function(a1, a2) {
  measure_distance(c(a1, a2), sim1)
}

models <- models %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))
models
#> # A tibble: 250 x 3
#>       a1      a2  dist
#>    <dbl>   <dbl> <dbl>
#> 1 -15.2   0.0889  30.8
#> 2  30.1  -0.827   13.2
#> 3  16.0   2.27    13.2
#> 4 -10.6   1.38    18.7
#> 5 -19.6  -1.04    41.8
#> 6   7.98  4.59    19.3
#> # … with 244 more rows
```
```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(models, rank(dist) <= 10)
  )
```

```{r}
ggplot(models, aes(a1, a2)) +
  geom_point(data = filter(models, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist))
```

or be more systematic
```{r}
grid <- expand.grid(
  a1 = seq(-5, 20, length = 25),
  a2 = seq(1, 3, length = 25)
  ) %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))

grid %>% 
  ggplot(aes(a1, a2)) +
  geom_point(data = filter(grid, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist)) 
```
```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(grid, rank(dist) <= 10)
  )
```
```{r}
best <- optim(c(0, 0), measure_distance, data = sim1)
best$par
#> [1] 4.222248 2.051204

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])
```
```{r}
sim1_mod <- lm(y ~ x, data = sim1)
coef(sim1_mod)
#> (Intercept)           x 
#>    4.220822    2.051533
```

Exercises:
```{r}
sim1a <- tibble(
  x = rep(1:10, each = 3),
  y = x * 1.5 + 6 + rt(length(x), df = 2)
)

sim1a_mod <- lm(y ~ x, data = sim1a)

ggplot(sim1a,aes(x,y)) +
  geom_point() +
   geom_abline(intercept = sim1a_mod$coefficients[1], 
               slope = sim1a_mod$coefficients[2], colour = "grey40")


sim1b <- tibble(
  x = rep(1:10, each = 3),
  y = x * 1.5 + 6 + rt(length(x), df = 2)
)

sim1b_mod <- lm(y ~ x, data = sim1b)

ggplot(sim1b,aes(x,y)) +
  geom_point() +
   geom_abline(intercept = sim1b_mod$coefficients[1], 
               slope = sim1b_mod$coefficients[2], colour = "grey40")
```

Predictions
```{r}
## buid some data
grid <- sim1 %>% 
  data_grid(x) 
## add predictions
grid <- grid %>% 
  add_predictions(sim1_mod) 

## viusalize
ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred), data = grid, colour = "red", size = 1)
```
Residuals
```{r}
sim1 <- sim1 %>% 
  add_residuals(sim1_mod)

## distribution
ggplot(sim1,aes(resid)) +
  geom_freqpoly(binwidth=0.5)
## based on predictor
ggplot(sim1,aes(x,resid)) +
  geom_point() +
  geom_ref_line(h = 0)
```
Exercises
```{r}


ggplot(data=sim1,aes(x,y)) +
  geom_point() +
  geom_smooth(method = )

lo <- loess(y~x,data = sim1)
```
```{r}
df <- tribble(
  ~y, ~x1, ~x2,
  4, 2, 5,
  5, 1, 6
)

model_matrix(df, y ~ x1)
```
```{r}
ggplot(sim2) + 
  geom_point(aes(x, y))

mod2 <- lm(y ~ x, data = sim2)
grid <- sim2 %>% 
  data_grid(x) %>% 
  add_predictions(mod2)
grid
```
```{r}
ggplot(sim2, aes(x)) + 
  geom_point(aes(y = y)) +
  geom_point(data = grid, aes(y = pred), colour = "red", size = 4)
```
```{r}
ggplot(sim3, aes(x1, y)) + 
  geom_point(aes(colour = x2))
```
```{r}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)

grid <- sim3 %>% 
  data_grid(x1, x2) %>% 
  gather_predictions(mod1, mod2)
grid
```
```{r}
ggplot(sim3, aes(x1, y, colour = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~ model)
```
```{r}
sim3 <- sim3 %>% 
  gather_residuals(mod1, mod2)

ggplot(sim3, aes(x1, resid, colour = x2)) + 
  geom_point() + 
  facet_grid(model ~ x2)
```

